---
title: "Agriculture"
author: "Brian Colgrove"
date: "4/8/2021"
output: html_document
---


```{r}
#packages
library(gstat)
library(tidyverse)
library(corrplot)
library(lmtest)
library(multcomp)
library(nlme)
library(DataExplorer)
library(sf)
library(maps)
library(mapproj)
library(ggmap)
library(car)
source("stdres.R")
source("predictgls.R")


#read in the data
agriculture <- read.table("https://mheaton.byu.edu/docs/files/Stat469/Topics/3%20-%20SpatialCorrelation/1%20-%20PointReference/HWCaseStudy/Data/WaterHoldingCapacity.txt", header = TRUE, sep = " ")

#summarize the data
head(agriculture)
summary(agriculture)
```


```{r}
#scatter plot Yield vs WHC
ggplot(agriculture, aes(x = Yield, y = WHC)) +
  geom_point()

#scatter plot EC vs WHC
ggplot(agriculture, aes(x = EC, y = WHC)) +
  geom_point()
```
There seems to be a positive linear relationship between Yield and WHC and also EC and WHC. 


```{r}
#lm model
agriculture.lm <- lm(WHC~EC + Yield, data = agriculture)
plot(agriculture.lm$fitted, agriculture.lm$residuals)

#variogram
data_omit <- na.omit(agriculture)
myVariogram <- variogram(object=WHC~EC + Yield, locations=~Lon+Lat, data=data_omit)
plot(myVariogram)

#residual plot
ggplot() + geom_tile(aes(x=data_omit$Lon, y=data_omit$Lat, fill=agriculture.lm$residuals)) + scale_fill_distiller(palette="Spectral")
```
There seems to be spatial correlation for both residual and variogram graphs. The residual one gets higher values the further North East you go. The variogram shows smaller variance at smaller distance. 


```{r}
#gls model
agriculture.gls <- gls(model=WHC~EC+Yield, data=data_omit, correlation=corExp(form=~Lon+Lat, nugget=TRUE), method="ML", na.action = na.omit)

agriculture.gls2 <- gls(WHC~Yield+EC, data=data_omit, correlation=corSpher(form=~Lon+Lat, nugget=TRUE), method="ML")

agriculture.gls3 <- gls(WHC~Yield+EC, data=data_omit, correlation=corGaus(form=~Lon+Lat, nugget=TRUE), method="ML")

c(AIC(agriculture.gls), AIC(agriculture.gls2), AIC(agriculture.gls3))



```
The exponential correlation gls model is just barely better than the other two according to its AIC values. 


```{r}
#linearity
avPlots(agriculture.lm)

#decorrelated residuals
decorResids <- stdres.gls(agriculture.gls)

#variogram of decorrelated residuals
residDF <- data.frame(Lon=data_omit$Lon, Lat=data_omit$Lat, decorrResid=decorResids)
residVariogram <- variogram(object=decorrResid~1, locations=~Lon+Lat, data=residDF)
plot(residVariogram)

#normality
hist(decorResids)

#residual plot
ggplot() + geom_tile(aes(x=data_omit$Lon, y=data_omit$Lat, fill=decorResids)) + scale_fill_distiller(palette="Spectral")
```
AV plots are linear, the histogram is normal, the variogram shows a more constant variance along the distances, the residual plot shows pretty good values across the map. No patterns. All assumptions are met.  


```{r}
#table
summary(agriculture.gls)$tTable

#confidence interval
intervals(agriculture.gls, which="coef")

#na predictions
vec <- which(is.na(agriculture$WHC))
na_data <- agriculture[vec,c(1:4) ]

#na.preds <- predict(temp.lm, newdata=na_data)
na.preds <- predictgls(agriculture.gls, newdframe = na_data)

#map predictions
colnames(na.preds)[5] <- "WHC"
full <- rbind(na.preds[1:5], data_omit )
mymap <- get_stamenmap(bbox = c(left = min(full$Lon), bottom = min(full$Lat), right =
  max(full$Lon), top =  max(full$Lat)), zoom=7, maptype="watercolor")
ggmap(mymap)+geom_raster(data=full, aes(x=Lon, y=Lat, fill=WHC),  alpha=0.8)+ 
  geom_tile()+ scale_fill_distiller(palette="Spectral", na.value=NA) + coord_cartesian()
```
From our hypothesis test it does seem like that higher yield areas increase with higher WHC since our p-value is 0.00778. We are 95% confident that WHC increases per 1 unit of Yield is between 0.006, 0.045. 

